---
title: "Example R Markdown Report"
author: "PennyPegman"
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: yes
---

```{r setup, include=FALSE, echo = FALSE}
source("~/alldbconnection.R")
connection <- alldbconnect("~/credentials.R", hive = TRUE, postgres = FALSE)

library(dplyr)
library(ggplot2) # for charts 
library(scales) # needed for percent labels
library(reshape2)
library(knitr) # required for table
library(kableExtra) # nice table formatting
library(lubridate)
library(DiagrammeR) #flowcharts
library(plotly)
library(magrittr)
library(sparklyr)
library(htmltools)
library(htmlwidgets)
library(reticulate)
library(webshot)
library(Hmisc)
library(DBI)
library(RODBC)
library(data.table)
library(DT)
library(readr)
library(RColorBrewer)
```


```{r load_data,  echo = FALSE, message=FALSE, error=FALSE, warning=FALSE}
system("hadoop fs -copyToLocal /user/burbankds/others/adhocs/LeedsFestival2019/age.csv ~/age.csv" ) 
system("hadoop fs -copyToLocal /user/burbankds/others/adhocs/LeedsFestival2019/LThealthdisability.csv ~/LThealthdisability.csv" ) 
system("hadoop fs -copyToLocal /user/burbankds/others/adhocs/LeedsFestival2019/ESAclaimants.csv ~/ESAclaimants.csv" ) 
system("hadoop fs -copyToLocal /user/burbankds/others/adhocs/LeedsFestival2019/claimantcount_leeds.csv ~/claimantcount_leeds.csv" ) 
system("hadoop fs -copyToLocal /user/burbankds/others/adhocs/LeedsFestival2019/qualifications.csv ~/qualifications.csv" ) 

# Read CSVs
age <- read_csv("~/age.csv")
healthdis <- read_csv("~/LThealthdisability.csv")
esa <- read_csv("~/ESAclaimants.csv")
claimantcount <- read_csv("~/claimantcount_leeds.csv")
qualifications <- read_csv("~/qualifications.csv")


# Rename columns
names(age)[1:23] = c("date","LA","geographyCode","RuralUrban","total","Age0to4","Age5to7","Age8to9","Age10to14","Age15","Age16to17","Age18to19","Age20to24",
                                "Age25to29","Age30to44","Age45to59", "Age60to64", "Age65to74", "Age75to84", "Age85to89", "Age90over", "MeanAge", "MedianAge")
names(healthdis)[1:8] = c("date","LA","geographyCode","RuralUrban","allhealthdis","limitedlot","limitedlittle","limitednot")

names(qualifications)[1:17] = c("date","LA","geogcode","ruralurban","All","No","Level1","Level2","Apprenticeships","Level3","Level4","Other",
                                "Students16to17","Students18over","StudentsEconActive_employment", "StudentsEconActive_Unemp", "StudentsEconInactive")


# add code in here to remove CSVs from user home area so ok if re-run it
```

```{r format_data, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE}

#  AGE: condense down age bands and select only variables needed

age_bands <- age %>% 
       mutate(Age0to15 = Age0to4 + Age5to7 + Age8to9 + Age10to14 + Age15) %>%
       mutate(Age18to24 = Age18to19 + Age20to24 ) %>%
       mutate(Age25to59 = Age25to29 + Age30to44 + Age45to59) %>%
       mutate(Age65over = Age65to74 + Age75to84 + Age85to89 + Age90over) %>%
       select(LA, total, Age0to15, Age16to17, Age18to24, Age25to59, Age60to64, Age65over)
# Add percentage in each age band
age_bands <- age_bands %>%
              mutate(age0to15_pct = (Age0to15 / total *100)) %>%
              mutate(age16to17_pct = (Age16to17 / total *100)) %>%
              mutate(age18to24_pct = (Age18to24 / total *100)) %>%
              mutate(age25to59_pct = (Age25to59 / total *100)) %>%
              mutate(age60to64_pct = (Age60to64 / total *100)) %>%
              mutate(age65over_pct = (Age65over / total *100))

# Restrict to just 5 LAs for comparision
age_bands_5LAs <- age_bands %>%
                  subset(LA %in% c("Leeds","Harrogate","York","Sheffield","Kingston upon Hull, City of"))
# Rename Hull
age_bands_5LAs$LA[age_bands_5LAs$LA == "Kingston upon Hull, City of"] = "Hull"

# End of Age

# Health Problems



# Socio economic Status


# Qualifications


# ESA claimants over time
esa$Date <- as.Date(esa$Date, "%m/%d/%Y")

# Claimant count


# Add a Publication date
pub_date <- (format(today(),"%d %B %Y")) # Changed so it automatically updates the date of the report

```

Published: <span class="highlightGrey">`r pub_date`</span>
<br><br>

R markdown has many benefits:
* You can embed explanatory text with charts, tables and other visuals, 
* Include downloadable data links
* Structure your report and include a table of contents on the left for easy navigation
* Save as static HTML and send to your customers or host on a webpage
* You can also output as slides, PDF and other formats
* When the data changes you just import the new data and everything will update automatically - you can even embed "r variables" into the text sections so it will update the text as well as the data changes - e.g the published date above is automatic and changes based on when you run it (date is an r variable that gets updated)

You **don't** need to be experienced in developing web pages as it's all very simple to code and put together.
There are cheatsheets and easy to follow documentation to help you.

However if you do have experience you can improve the presentation further by including CSS Style sheets and other features.

This report shows some examples of charts which can be used with R markdown to communicate insights from data.
 
## Total Population 

As at the Census 2011

Need to get the labels sorted so right at the end and make the pretty number
Also get the bars in descending order

```{r populationchart, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE}

library(plotly)


y <- age_bands_5LAs$LA
x1 <- round(age_bands_5LAs$total,0)

# Convert the Site name to proper case
proper <- function(x) paste0(toupper(substr(x, 1, 1)), tolower(substring(x, 2)))
y <- proper(y)

population <- data.frame(y, x1)

# Arrange in ascending order so will be in descending order on chart
population = population %>%
            arrange(x1)

# Generate the plot_ly chart
plot_ly(population, x = ~x1, name='Total', y = ~y, type = 'bar', orientation = 'h',
        marker = list(color = 'steelblue',
                      line = list(color = 'white', width = 1))) %>%
  
  layout(xaxis = list(title = "",
                      showgrid = FALSE,
                      showline = FALSE,
                      showticklabels = FALSE,
                      zeroline = FALSE,
                      domain = c(0.15, 1)),
         yaxis = list(title = "",
                      showgrid = FALSE,
                      showline = FALSE,
                      showticklabels = FALSE,
                      zeroline = FALSE),
         barmode = 'stack',
         paper_bgcolor = 'white', plot_bgcolor = 'white',
         margin = list(l = 10, r = 10, t = 10, b = 10),
         showlegend = FALSE) %>%
  # labeling the y-axis
  add_annotations(xref = 'paper', yref = 'y', x = 0.13, y = y,
                  xanchor = 'right',
                  text = paste('<b>',y,'</b>'),
                  font = list(family = 'Arial', size = 12,
                            color = 'rgb(67, 67, 67)'),
                  showarrow = FALSE, align = 'right') %>%
  # labeling the percentages of each bar (x_axis)
  add_annotations(xref = 'x', yref = 'y',
                  x = x1 + 1500, y = y,
                  text = paste('<b>',population[,"x1"]),
                  font = list(family = 'Arial', size = 12,
                            color = 'grey'),
                  showarrow = FALSE) 

#age_bands_5LAs_sorted = age_bands_5LAs %>%
 #                           arrange(desc(total))
# Bar Chart showing number of attempts by Bank

# 
# ggplot(data=age_bands_5LAs_sorted, aes(x=reorder(LA, total), y=total)) +
#   geom_bar(stat="identity", position=position_dodge(), fill='steelblue') +
#   geom_text(aes(label = population), hjust=-0.1, size=3, position=position_dodge(width = 1), inherit.aes = TRUE) +
#   ylab ('Population') +
#   xlab ('LA') +
#   theme(legend.title=element_blank()) +
#   ggtitle('Population') +
#   coord_flip()


```



## Age Distribution

You can use colour-blind friendly palettes for your data visualisations see: <https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/>

This chart shows the age distribution of five local authorities:
* this bar chart can enable us to compare how similar the age distribution of **Leeds** is to other local authorities in the area.  
* This is taken from Census 2011 data.

** PERHAPS CHANGE COLOUR SCHEME TO MATCH THE OTHER CHARTS ON THIS REPORT **

```{r agedistchart, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE}

#display.brewer.all(colorblindFriendly = TRUE)
# Hexadecimal color specification 
#brewer.pal(n = 6, name = "Set2")

# Chart of Age distribution
y <- age_bands_5LAs$LA
x1 <- round(age_bands_5LAs$age0to15_pct,0)
x2 <- round(age_bands_5LAs$age16to17_pct,0)
x3 <- round(age_bands_5LAs$age18to24_pct,0)
x4 <- round(age_bands_5LAs$age25to59_pct,0)
x5 <- round(age_bands_5LAs$age60to64_pct,0)
x6 <- round(age_bands_5LAs$age65over_pct,0)

age_dist <- data.frame(y, x1, x2, x3, x4, x5, x6)


top_labels <- c('Under 16', '16 to 17', '18 to 24', '25 to 59', '60 to 64', '65 and over')


plot_ly(age_dist, x = ~x1, name='Under 16', y = ~y, type = 'bar', orientation = 'h',
        marker = list(color = '#1B9E77',
                      line = list(color = 'white', width = 1))) %>%
  add_trace(x = ~x2, name='16 to 17', marker = list(color = '#D95F02')) %>%
  add_trace(x = ~x3, name='18 to 24', marker = list(color = '#7570B3')) %>%
  add_trace(x = ~x4, name='25 to 59', marker = list(color = '#E7298A')) %>%
  add_trace(x = ~x5, name='60 to 64', marker = list(color = '#66A61E')) %>%
  add_trace(x = ~x6, name='65 and over', marker = list(color = '#E6AB02')) %>%
  
#"#1B9E77" "#D95F02" "#7570B3" "#E7298A" "#66A61E" "#E6AB02" "#A6761D"
  layout(xaxis = list(title = "",
                      showgrid = FALSE,
                      showline = FALSE,
                      showticklabels = FALSE,
                      zeroline = FALSE,
                      domain = c(0.15, 1)),
         yaxis = list(title = "",
                      showgrid = FALSE,
                      showline = FALSE,
                      showticklabels = FALSE,
                      zeroline = FALSE),
         barmode = 'stack',
         paper_bgcolor = 'white', plot_bgcolor = 'white',
         margin = list(l = 10, r = 10, t = 10, b = 10),
         showlegend = TRUE) %>%
  
  # labeling the y-axis
  add_annotations(xref = 'paper', yref = 'y', x = 0.14, y = y,
                  xanchor = 'right',
                  text = paste('<b>',y,'</b>'),
                  font = list(family = 'Arial', size = 12,
                            color = 'rgb(67, 67, 67)'),
                  showarrow = FALSE, align = 'right') %>%
  # labeling the percentages of each bar (x_axis)
  add_annotations(xref = 'x', yref = 'y',
                  x = x1 / 2, y = y,
                  text = paste(age_dist[,"x1"], '%'),
                  font = list(family = 'Arial', size = 12,
                            color = 'white'),
                  showarrow = FALSE) %>%
  add_annotations(xref = 'x', yref = 'y',
                  x = x1 +x2/2, y = y,
                  text = paste(age_dist[,"x2"], '%'),
                  font = list(family = 'Arial', size = 12,
                              color = 'white'),
                  showarrow = FALSE) %>%
  add_annotations(xref = 'x', yref = 'y',
                  x = x1 + x2 + x3 / 2, y = y,
                  text = paste(age_dist[,"x3"], '%'),
                  font = list(family = 'Arial', size = 12,
                              color = 'white'),
                  showarrow = FALSE) %>%
  add_annotations(xref = 'x', yref = 'y',
                  x = x1 + x2 + x3 +x4 / 2, y = y,
                  text = paste(age_dist[,"x4"], '%'),
                  font = list(family = 'Arial', size = 12,
                              color = 'white'),
                  showarrow = FALSE) %>%
    add_annotations(xref = 'x', yref = 'y',
                  x = x1 + x2 + x3 +x4 + x5 / 2, y = y,
                  text = paste(age_dist[,"x5"], '%'),
                  font = list(family = 'Arial', size = 12,
                              color = 'white'),
                  showarrow = FALSE) %>%
    add_annotations(xref = 'x', yref = 'y',
                  x = x1 + x2 + x3 +x4 + x5 +x6 / 2, y = y,
                  text = paste(age_dist[,"x6"], '%'),
                  font = list(family = 'Arial', size = 12,
                              color = 'white'),
                  showarrow = FALSE)

```

## Health Problems and Disabilities

This shows...  This is taken from Census 2011 data.

```{r healthchart, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE}



```


## Socio Economic Status

This is taken from Census 2011 data

## Qualifications

This is taken from Census 2011 data.
There is also data on the level of the qualifications.

** ADD code to round the percentagesl and format the numbers in thousands** 

This table shows the percentage of the population aged 16 and over with No Qualifications and also those with apprenticeships as their highest qualification

```{r qualificationschart, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE}

qualifications2 <- qualifications %>%
                  select(LA, All, No, Level1, Level2, Apprenticeships, Level3, Level4, Other) %>%                   
                  subset(LA %in% c("Leeds","Harrogate","York","Sheffield","Kingston upon Hull, City of")) %>%
                  mutate(pct_noqual = percent(No/All)) %>%
                  mutate(pct_apprenticeship = percent(Apprenticeships/All))
# Rename Hull
qualifications2$LA[qualifications2$LA == "Kingston upon Hull, City of"] = "Hull"

kable(qualifications2[,c("LA","All","pct_noqual", "pct_apprenticeship")], format = "markdown", row.names = FALSE,  col.names = c("Local Authority","Total Population aged 16 and over", "% with No qualification", "% with Apprenticeships"))
```


## Benefit Claimants

This section looks at those claiming benefits - those on ESA (all 5 Local Authorities) and the claimant count (Leeds only)


### ESA Claimants over time

Need to change the colours of the lines and get the annotation for ESA introduction working

```{r esaplot, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE}
library(plotly)

x <- as.factor(esa$Date)
y_harrogate <- esa$Harrogate
y_hull <- esa$Hull
y_leeds <- esa$Leeds
y_sheffield <- esa$Sheffield
y_york <- esa$York

data <- data.frame(x, y_harrogate, y_hull, y_leeds, y_sheffield, y_york)

xaxis <- list(title = "",
               showticklabels = TRUE,
              showline = FALSE,
              tickangle = 315,
              exponentformat = "E",
              dtick=12,
             ticks = 'outside',
             tickcolor = 'rgb(204, 204, 204)',
             tickfont = list(family = 'Arial',
                             size = 10,
                             color = 'rgb(82, 82, 82)'))




yaxis <- list(title = "",
             showline = FALSE,
             showgrid = TRUE,
             showticklabels = TRUE,
             linecolor = 'rgb(204, 204, 204)',
             linewidth = 2,
             autotick = TRUE,
             ticks = 'outside',
             tickcolor = 'rgb(204, 204, 204)',
             tickwidth = 2,
             ticklen = 5,
             tickfont = list(family = 'Arial',
                             size = 10,
                             color = 'rgb(82, 82, 82)'))

margin <- list(autoexpand = FALSE,
              l = 100,
              r = 100,
              t = 110)

# Build the annotations

rise_annotation <- list(
  xref = 'paper',
  yref = 'y',
  x = "2011-01-08",
  y = 8000,
 #xanchor = 'right',
#  yanchor = 'bottom',
  text = ~paste('ESA was being introduced'),
  arrowhead = 4,
  arrowsize = .5,
  arrowcolor="#CCCCCC",
  ax = 0,
  ay = -40,
  font = list(family = 'Arial',
              size = 10,
              color = '#AAAAAA'),
  showarrow = TRUE)

harrogate_annotation <- list(
  xref = 'paper',
  x = 1,
  y = y_harrogate[length(y_harrogate)],
  xanchor = 'left',
  yanchor = 'middle',
  text = paste(prettyNum(y_harrogate[length(y_harrogate)],big.mark=",")),
    font = list(family = 'Arial',
                size = 12,
                color = 'grey'),
  showarrow = FALSE)

hull_annotation <- list(
  xref = 'paper',
  x = 1,
  y = y_hull[length(y_hull)]+500,
  xanchor = 'left',
  yanchor = 'middle',
  text = paste(prettyNum(y_hull[length(y_hull)],big.mark=",")),
    font = list(family = 'Arial',
                size = 12,
                color = 'hsl(33, 100%, 45%)'),
  showarrow = FALSE)

leeds_annotation <- list(
  xref = 'paper',
  x = 1,
  y = y_leeds[length(y_leeds)]-500,
  xanchor = 'left',
  yanchor = 'middle',
  text = paste(prettyNum(y_leeds[length(y_leeds)],big.mark=",")),
    font = list(family = 'Arial',
                size = 12,
                color = 'hsl(207, 44%, 55%)'),
  showarrow = FALSE)


sheffield_annotation <- list(
  xref = 'paper',
  x = 1,
  y = y_sheffield[length(y_sheffield)]-500,
  xanchor = 'left',
  yanchor = 'middle',
  text = paste(prettyNum(y_sheffield[length(y_sheffield)],big.mark=",")),
    font = list(family = 'Arial',
                size = 12,
                color = 'hsl(207, 44%, 55%)'),
  showarrow = FALSE)

york_annotation <- list(
  xref = 'paper',
  x = 1,
  y = y_york[length(y_york)]-500,
  xanchor = 'left',
  yanchor = 'middle',
  text = paste(prettyNum(y_york[length(y_york)],big.mark=",")),
    font = list(family = 'Arial',
                size = 12,
                color = 'hsl(207, 44%, 55%)'),
  showarrow = FALSE)




plot_ly(data, x = ~x) %>%
  add_trace(y = ~y_harrogate, type = 'scatter', name = "Harrogate", mode = 'lines', line = list(color = '#666666', width = 2))  %>%
  add_trace(y = ~y_hull, type = 'scatter', name = "Hull", mode = 'lines', line = list(color = 'hsl(33, 100%, 45%)', width = 2)) %>%
  add_trace(y = ~y_leeds, type = 'scatter', name = "Leeds", mode = 'lines', line = list(color = 'hsl(207, 44%, 55%)', width = 2)) %>%
    add_trace(y = ~y_sheffield, type = 'scatter', name = "Sheffield", mode = 'lines', line = list(color = 'hsl(207, 44%, 55%)', width = 2)) %>%
    add_trace(y = ~y_york, type = 'scatter', name = "York", mode = 'lines', line = list(color = 'hsl(207, 44%, 55%)', width = 2)) %>%
  layout(title = "ESA claimants over time", xaxis = xaxis, yaxis = yaxis, margin = margin,
         autosize = FALSE,
         showlegend = TRUE,
           annotations = rise_annotation) %>%
          layout(annotations = harrogate_annotation)  %>%
          layout(annotations = hull_annotation)  %>%
          layout(annotations = leeds_annotation) %>%
          layout(annotations = sheffield_annotation) %>%
          layout(annotations = york_annotation) %>%
          layout(legend = list(x = 0.05, y = 0.9)) %>%
          layout(margin = list(l = 50, 
                               r = 50, 
                               b = 90,
                               t = 70,
                               pad = 0)) %>% # Stops the x labels getting cut off
  layout(width = 780) %>%
  layout(height = 460)


```

### Claimant Count (Leeds only)

A Time-series of the Claimant count in Leeds, by gender

```{r claimantcountplot, echo = FALSE, message=FALSE, error=FALSE, warning=FALSE}

x <- as.Date(claimantcount$Month, "%m/%d/%Y")
y_total <- claimantcount$Total
y_males <- claimantcount$Male
y_females <- claimantcount$Female

claimantcount_data <- data.frame(x, y_total, y_males, y_females) %>%
  mutate(formatted_date = as.Date(x, format= "%Y-%m-%d"))

xaxis <- list(title = "",
             linecolor = 'black',
             linewidth = 2,
             mirror = TRUE,
              showticklabels = TRUE,
              showline = FALSE,
              tickangle = 315,
              exponentformat = "E",
              dtick=1,
              ticks = 'outside',
              tickcolor = 'rgb(204, 204, 204)',
              tickfont = list(family = 'Arial',
                             size = 12,
                             color = 'rgb(82, 82, 82)'))

yaxis <- list(title = "",
             linecolor = 'black',
             linewidth = 2,
             mirror = TRUE,
             showline = FALSE,
             showgrid = TRUE,
             showticklabels = TRUE,
             linecolor = 'rgb(204, 204, 204)',
             linewidth = 2,
             autotick = TRUE,
             ticks = 'outside',
             tickcolor = 'rgb(204, 204, 204)',
             tickwidth = 0.5,
             ticklen = 5,
             tickfont = list(family = 'Arial',
                             size = 12,
                             color = 'rgb(82, 82, 82)'))

num_decs_annotation <- list(
  xref = 'paper',
  x = 0.965,
  y = y_total[length(y_total)]+1000,
  xanchor = 'left',
  yanchor = 'middle',
  text = paste("Total declarations:<br>",prettyNum(y_total[length(y_total)],big.mark=",")),
    font = list(family = 'Arial',
                size = 12,
                color = 'rgba(55, 128, 191'),
  showarrow = FALSE,
  bgcolor = "#E2E2E2",
  bordercolor = "#444",
  borderwidth = 1)

l <- list(
  font = list(
    family = "sans-serif",
    size = 12,
    color = "#000"),
  bgcolor = "#E2E2E2",
  bordercolor = "#444",
  borderwidth = 1,
  x = 0.05,
  y = 1.05)

t <- list(
  family = "Arial",
  size = 18,
  color = 'black')

plot_ly(claimantcount_data, x = ~as.factor(x)) %>%
  add_trace(y = ~y_males, type = 'bar', name = "Males", 
            marker = list(color = 'rgba(55, 128, 191, 0.65)', line = list(color = 'rgba(55, 128, 191, 0.7)', width = 2))) %>%
    add_trace(y = ~y_females, type = 'bar', name = "Females", 
              marker = list(color = 'rgba(255, 165, 0, 0.65)', line = list(color = 'rgba(255, 165, 0, 0.7)', width = 2))) %>%
              
  layout(title = "", font = t, xaxis = xaxis, yaxis = yaxis, margin = margin, barmode = 'stack',
         autosize = TRUE,
         showlegend = TRUE) %>%
          layout(margin = list(l = 50, 
                               r = 50, 
                               b = 90,
                               t = 70,
                               pad = 0)) %>% # Stops the x labels getting cut off
          layout(legend = l) %>%
  layout(width = 780) %>%
  layout(height = 460)
```


## Download Data


<div RawDownloadLinksBox id="RawDownloadLinksBox">
Download the ESA data <span class="highlightBlue">weekly figures:</span> 

```{r data0,echo=FALSE,  message=FALSE, error=FALSE, warning=FALSE}



DT::datatable(esa, 
              extensions = 'Buttons',   options = list(
                      bPaginate = FALSE, dom = 'Bfrtip',
                      buttons =
                        list(list(
                          extend = 'collection', 
                          buttons = list(c(extend = 'csv', filename= 'esa_data'),c(extend = 'excel', filename= 'esa_data')),
                          text = 'Download')) )) %>%
                        onRender("
                            function dropTable0() {
    document.getElementById('DataTables_Table_0').style.display = 'none';
    document.getElementById('DataTables_Table_0_info').style.display = 'none';
    document.getElementById('DataTables_Table_0_filter').style.display = 'none';}
                                 ")


```
</div>


```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE, results="hide"}


# remove CSV once converted to dataframe and no longer needed in any other code sections
file.remove("~/claimantcount_leeds.csv") # will state TRUE as output when file is deleted
file.remove("~/ESAclaimants.csv") # will state TRUE as output when file is deleted
file.remove("~/LThealthdisability.csv") # will state TRUE as output when file is deleted
file.remove("~/age.csv") # will state TRUE as output when file is deleted
file.remove("~/qualifications.csv") # will state TRUE as output when file is deleted
#file.remove("~/claimantcount_leeds.csv") # will state TRUE as output when file is deleted

```

